---
title: "Cookie Cats A/B Testing"
format: html
editor: visual
output-file: index.html
---

## 1. Define the experiment and research question

Cookie Cats is a mobile puzzle game developed by Tactile Entertainment.

As players progress through the game, they encounter gates that require them to either wait before continuing or make an in-app purchase to continue immediately.

In this **experiment**, players were randomly assigned to one of two versions of the game:

- The first gate placed at level 30

- The first gate placed at level 40.

Formally, the **research question** is to determine whether moving the first gate from level 30 to 40 helps improve player engagement and retention.

Reading the data:

```{r}
data = read.csv("cookie_cats.csv")
```

There are five variables:

- **userid** - (int) a unique number that identifies each player.

- **version** - (gate_30/ gate_40) the version of the game assigned to the player:

  - gate_30: the first gate is at level 30.

  - gate_40: the first gate is at level 40.

- **sum_gamerounds** - (int) the number of game rounds played by the player during the first 14 days after install.

- **retention_1** - (True/False) did the player come back and play 1 day after installing?

- **retention_7** - (True/False) did the player come back and play 7 days after installing?

Since we want to know "whether moving the first gate from level 30 to 40 helps improve player engagement and retention", we will be examining the effect of **version** (explanatory) on the **retention** and **number of game rounds** (response).

## 2. Identify the treatment and control groups

The players were randomly assigned to one of two versions of the game

- **Control group**: gate_30

- **Treatment group**: gate_40

```{r, echo = FALSE}
table(data$version)
```

Roughly the same amount of players were assigned to each group.

## 3. Define the relevant experiment metrics

The outcomes available in the dataset measure two aspects of player behavior:

- **1-day retention**: whether a player returns 1 day after installing the game.

- **7-day retention**: whether a player returns 7 days after installing the game.

- **Game rounds**: the total number of rounds played during the first 14 days.

We will use these metrics to compare player behavior between the control and treatment groups.

## 4. Explore and check the experimental data

Before estimating the treatment effect, examine the quality and the structure of the data.

**Dimension of the data**

```{r}
dim(data)
```

The data contains 90189 rows with 5 variables.

**Missing values**

```{r}
colSums(is.na(data))
```

None of the values are missing.

**Duplicate players**

```{r}
sum(duplicated(data$userid))
```

No duplicated players

**Variables**

1.  Retention

    ```{r}
    table(data$retention_1)
    table(data$retention_7)
    ```

    Roughly about half of the users returned on day 1, but most most users did not return on day 7.

    78.6% of users who returned on day 7 also returned on day 1.

    ```{r}
    sum(
      data$retention_1 == "True" &
        data$retention_7 == "True"
    )/sum(
        data$retention_7 == "True"
    )
    ```

2.  Game Rounds

    ```{r}
    summary(data$sum_gamerounds)
    hist(data$sum_gamerounds)
    ```

    There seem to be some substantial skewness in the game rounds. The mean is much higher at 52 compared to the median at 16, implying an extreme right-skewness: while the middle 50% of the users played between 5\~50 rounds, a small number of players logged 2,000\~50,000 rounds.

    ```{r}
    head(data[order(data$sum_gamerounds, decreasing = TRUE), ], 5)
    ```

    Because this extreme value makes the raw distribution difficult to visualize, we can examine the distribution on a log scale:

    ```{r}
    hist(log(data$sum_gamerounds), 
         breaks = 50, 
         main = "Distribution of Game Rounds", 
         xlab = "log(Game Rounds)" )
    ```

    Instead of removing the extreme observations, we will later examine whether they substantially influence the estimated treatment effect.

## 5. Estimate and test the treatment effect

Because players were randomly assigned to gate 30 or gate 40, we can estimate the treatment effect by directly comparing outcomes between the two groups. Throughout the analysis:

- A = gate 30 (control)

- B = gate 40 (treatment)

For each outcome, the treatment effect is defined as:

$\Delta = \text{Outcome}_B - \text{Outcome}_A$

A positive value of $\Delta$ favors gate 40, while a negative value favors gate 30.

First, convert the retention variables to 0/1 indicators:

```{r}
data$retention_1_num <- as.integer(
  tolower(as.character(data$retention_1)) == "true"
)

data$retention_7_num <- as.integer(
  tolower(as.character(data$retention_7)) == "true"
)

n_A <- sum(data$version == "gate_30")
n_B <- sum(data$version == "gate_40")
```

### 5.1 1-day Retention

For 1-day retention, let $p_A$ be the probability that a player assigned to gate 30 returns after one day, and let $p_B$ be the corresponding probability for gate 40.

The treatment effect is $$\Delta_1 = p_B - p_A$$

We estimate these probabilities using the observed proportions in each group.

```{r}
p_A_1 <- mean( data$retention_1_num[data$version == "gate_30"] ) 
p_B_1 <- mean( data$retention_1_num[data$version == "gate_40"] ) 
delta_1 <- p_B_1 - p_A_1 
c( control_rate = p_A_1, treatment_rate = p_B_1, treatment_effect = delta_1 )
```

The hypotheses are:

$$H_0: p_B - p_A = 0 \\
H_A: p_B - p_A \neq 0$$

The standard error of the estimated difference in proportions is:

```{r}
se_1 <- sqrt( p_A_1 * (1 - p_A_1) / n_A + p_B_1 * (1 - p_B_1) / n_B ) 
se_1
```

An approximate 95% confidence interval for the treatment effect is:

```{r}
ci_1 <- delta_1 + c(-1, 1) * qnorm(0.975) * se_1 
ci_1
```

Finally, conduct a two-sample proportion test.

```{r}
test_1 <- prop.test(x = c( sum(data$retention_1_num[data$version == "gate_40"]), 
                           sum(data$retention_1_num[data$version == "gate_30"]) ),
                    n = c(n_B, n_A), correct = FALSE ) 
test_1
```

Here, gate 40 is entered first and gate 30 second so that the estimated difference corresponds to treatment minus control.

### 5.2 7-day Retention

For 7-day retention, let

$p_A = P(\text{retained at day 7} \mid \text{gate 30})$ and $p_B = P(\text{retained at day 7} \mid \text{gate 40})$,

where gate 30 is the control group and gate 40 is the treatment group.

The treatment effect of interest is the difference in 7-day retention probabilities: $\Delta = p_B - p_A$

Therefore:

- $\Delta >0$: moving the gate to level 40 increases 7-day retention.

- $\Delta < 0$: moving the gate to level 40 decreases 7-day retention.

- $\Delta = 0$: moving the gate has no effect on 7-day retention.

Because the true population probabilities $p_A$ and $p_B$are unknown, we will estimate them using the observed retention proportions in the two groups.

```{r}
p_A_7 <- mean(
  data$retention_7_num[data$version == "gate_30"]
)

p_B_7 <- mean(
  data$retention_7_num[data$version == "gate_40"]
)

delta_7 <- p_B_7 - p_A_7

c(
  control_rate = p_A_7,
  treatment_rate = p_B_7,
  treatment_effect = delta_7
)
```

The hypotheses are (same as day-1 retention):

$$H_0: p_B - p_A = 0 \\
H_A: p_B - p_A \neq 0$$

Calculate the standard error:

```{r}
se_7 <- sqrt(
  p_A_7 * (1 - p_A_7) / n_A +
    p_B_7 * (1 - p_B_7) / n_B
)

se_7
```

Construct an approximate 95% confidence interval:

```{r}
ci_7 <- delta_7 +
  c(-1, 1) * qnorm(0.975) * se_7

ci_7
```

Then perform the two-sample proportion test:

```{r}
test_7 <- prop.test(x = c( sum(data$retention_7_num[data$version == "gate_40"]), 
                           sum(data$retention_7_num[data$version == "gate_30"]) ), 
                    n = c(n_B, n_A), correct = FALSE ) 
test_7
```

### 5.3 Number of game rounds

The number of game rounds is a quantitative outcome rather than a binary outcome.

Let $\mu_A$ represent the mean number of game rounds for players assigned to gate 30 and $\mu_B$ represent the mean for players assigned to gate 40.

The treatment effect is:

$$\Delta_{\text{rounds}} = \mu_B - \mu_A$$

Estimate the group means and treatment effect:

```{r}
mean_A <- mean(
  data$sum_gamerounds[data$version == "gate_30"]
)

mean_B <- mean(
  data$sum_gamerounds[data$version == "gate_40"]
)

delta_rounds <- mean_B - mean_A

c(
  control_mean = mean_A,
  treatment_mean = mean_B,
  treatment_effect = delta_rounds
)
```

The hypotheses are:

$$H_0: \mu_B - \mu_A = 0 \\
H_A: \mu_B - \mu_A \neq 0$$

Since the population variances are unknown and must be estimated from the sample, we can use a two sample t-test for unequal variance (aka. Welch two-sample t-test).

The main difference between Welch's t-test and the equal-variance version of the two-sample t-test is the assumption about the two population variances:

- The equal-variance t-test assumes that both groups have the same population variance.

- Welch's t-test estimates the variance separately for each group and does not assume that the variances are equal.

Both tests compare the difference between two sample means with the uncertainty in that difference. Welch's test is generally the safer default because it remains appropriate when the groups have different variances or different sample sizes. It is not a z-test, because the population variances are unknown and must be estimated from the data.

In practice, this is done by calling a standard t.test in R with `var.equal = FALSE`:

```{r}
rounds_test <- t.test(
  data$sum_gamerounds[data$version == "gate_40"],
  data$sum_gamerounds[data$version == "gate_30"],
  alternative = "two.sided",
  var.equal = FALSE
)
rounds_test
```

The confidence interval reported by `t.test()` corresponds to the difference, $\mu_B - \mu_A$, where $\mu_B$ is the mean number of game rounds for gate 40 and $\mu_A$ is the mean number of game rounds for gate 30.

Because the distribution of game rounds is extremely right-skewed and contains an unusually large observation, the difference in means should be interpreted with some caution. The primary analysis will retain all observations, but we can also conduct a sensitivity analysis to determine how strongly the extreme observation affects the result.

## 6. Interpret the results

## 7. Make a practical recommendation
