---
title: "Cookie Cats A/B Testing"
format: html
editor: visual
---

## 1. Define the experiment and research question
Cookie Cats is a mobile puzzle game developed by Tactile Entertainment. 

As players progress through the game, they encounter gates that require them to either wait before continuing or make an in-app purchase to continue immediately. 

In this **experiment**, players were randomly assigned to one of two versions of the game: 
 * The first gate placed at level 30
 * The first gate placed at level 40. 
 
Formally, the **research question** is to determine whether moving the first gate from level 30 to 40 helps improve player engagement and retention. 


## 2. Identify the treatment and control groups

Now, reading the data: 

```{r}
data = read.csv("cookie_cats.csv")
```

There are five variables in the dataset: 
userid - a unique number that identifies each player.
version - whether the player was put in the control group (gate_30 - a gate at level 30) or the group with the moved gate (gate_40 - a gate at level 40).
sum_gamerounds - the number of game rounds played by the player during the first 14 days after install.
retention_1 - did the player come back and play 1 day after installing?
retention_7 - did the player come back and play 7 days after installing?

```{r, echo=FALSE}
str(data)
```


## 3. Define the relevant experiment metrics
## 4. Explore and check the experimental data
## 5. Estimate and test the treatment effect
## 6. Interpret the results
## 7. Make a practical recommendation