---
title: "Cookie Cats A/B Testing"
format: html
editor: visual
---

## 1. Define the experiment and research question

Cookie Cats is a mobile puzzle game developed by Tactile Entertainment.

As players progress through the game, they encounter gates that require them to either wait before continuing or make an in-app purchase to continue immediately.

In this **experiment**, players were randomly assigned to one of two versions of the game:

- The first gate placed at level 30

- The first gate placed at level 40.

Formally, the **research question** is to determine whether moving the first gate from level 30 to 40 helps improve player engagement and retention.

Reading the data:

```{r}
data = read.csv("cookie_cats.csv")
```

There are five variables:

- **userid** - (int) a unique number that identifies each player.

- **version** - (gate_30/ gate_40) the version of the game assigned to the player:

  - gate_30: the first gate is at level 30.

  - gate_40: the first gate is at level 40.

- **sum_gamerounds** - (int) the number of game rounds played by the player during the first 14 days after install.

- **retention_1** - (True/False) did the player come back and play 1 day after installing?

- **retention_7** - (True/False) did the player come back and play 7 days after installing?

Since we want to know "whether moving the first gate from level 30 to 40 helps improve player engagement and retention", we will be examining the effect of **version** (explanatory) on the **retention** and **number of game rounds** (response).

## 2. Identify the treatment and control groups

The players were randomly assigned to one of two versions of the game

- **Control group**: gate_30

- **Treatment group**: gate_40

```{r, echo = FALSE}
table(data$version)
```

Roughly the same amount of players were assigned to each group.

## 3. Define the relevant experiment metrics

The outcomes available in the dataset measure two aspects of player behavior:

- **1-day retention**: whether a player returns 1 day after installing the game.

- **7-day retention**: whether a player returns 7 days after installing the game.

- **Game rounds**: the total number of rounds played during the first 14 days.

We will use these metrics to compare player behavior between the control and treatment groups.

## 4. Explore and check the experimental data

Before estimating the treatment effect, examine the quality and the structure of the data.

**Dimension of the data**

```{r}
dim(data)
```

The data contains 90189 rows with 5 variables.

**Missing values**

```{r}
colSums(is.na(data))
```

None of the values are missing.

**Duplicate players**

```{r}
sum(duplicated(data$userid))
```

No duplicated players

**Variables**

1.  Retention

    ```{r}
    table(data$retention_1)
    table(data$retention_7)
    ```

    Roughly about half of the users returned on day 1, but most most users did not return on day 7.

    78.6% of users who returned on day 7 also returned on day 1.

    ```{r}
    sum(
      data$retention_1 == "True" &
        data$retention_7 == "True"
    )/sum(
        data$retention_7 == "True"
    )
    ```

2.  Game Rounds

    ```{r}
    summary(data$sum_gamerounds)
    hist(data$sum_gamerounds)
    ```

    There seem to be some substantial skewness in the game rounds. The mean is much higher at 52 compared to the median at 16, implying an extreme right-skewness: while the middle 50% of the users played between 5\~50 rounds, a small number of players logged 2,000\~50,000 rounds.

    ```{r}
    head(data[order(data$sum_gamerounds, decreasing = TRUE), ], 5)
    ```

    Because this extreme value makes the raw distribution difficult to visualize, we can examine the distribution on a log scale:

    ```{r}
    hist(log(data$sum_gamerounds), 
         breaks = 50, 
         main = "Distribution of Game Rounds", 
         xlab = "log(Game Rounds)" )
    ```

    Instead of removing the extreme observations, we will later examine whether they substantially influence the estimated treatment effect.

## 5. Estimate and test the treatment effect

## 6. Interpret the results

## 7. Make a practical recommendation
